Deep Learning

### 2015-01-07 2014/nature-review/deepreview.tex
updating bib entries and references cleanup

So far, most applications of deep learning use feed-forward neural
networks, but for tasks that involve sequential inputs it is often better
to use recurrent neural networks. These are given one element from an input % YB term-->element
sequence at each time step and their recurrent internal ``hidden'' state
depends on the whole history of the input sequence. Recurrent networks are
very good at predicting the next character in a string\cite{Sutskever} or
the next word in a sequence\cite{Mikolov} but they can also be used for
more complex tasks.  After reading an English sentence one word at a time,
for example, an English``encoder'' network can be trained so that the final
state vector of its hidden neurons is a good representation of the thought
expressed by the sentence.  This thought vector can then be used as the
initial hidden state of a trained French ``decoder'' network which outputs
a probability distribution for the first word of the French translation. If
a particular first word is chosen (e.g., according to that probability)
and provided as input to the decoder
network it will then output a probability distribution for the second word
of the translation and so on until a full stop is chosen~\cite{Vinyals,
  Montreal}. 
%YB added a clarification
Overall, this process generates sequences of French words according to
a probability distribution that depends on the English sentence.

@@

The hidden layers of a multi-layer neural network learn to represent the
network's inputs in a way that makes it easy to predict the target
outputs. This is nicely demonstrated by training a multi-layer neural
network to predict the next word in a sequence from a local context of
earlier words~\citep{BenDucVin01-short}. 
Each context word is presented to the network as a one-of-N
vector in which one component has a value of $1$ and the rest are $0$. The
output layer has one unit per word and is forced to produce a probability
distribution over words by using the ``softmax'' function:
\begin{equation}
p_j = \frac{exp(z_j)}{sum_i exp(z_i)}
\end{equation}
where $z_j$ is the total input received by output unit $j$. The hidden
layers use either the logistic or tanh non-linearity:
\begin{equation}
y_j = \frac{1}{1+ exp(z_j)}, \ \ \ y_j = \frac{}{} % YB: what was the intent here for the 2nd part?
\end{equation}

The connectivity of the network is restricted so that each of the one-of-N
input vectors must first be converted to a ``word-vector'' before being
used by subsequent layers to give a high probability to the target next
word.  The network learns word-vectors that contain many active components
each of which can be interpreted as a separate feature of the word. This
was first demonstrated\cite{RHW} using sequences of three words, such as
(Colin, Mother, Victoria), derived from two family trees. The learned word
vector for Colin contained ``semantic'' features representing which family
tree Colin was in, what branch of that tree he was in, and from what generation
he was~\cite{hinton86}. These semantic features were not explicitly present
in the input.  They were discovered by the learning procedure as a good way
of factorizing the structured relationships between the input and output
symbols into multiple ``micro-rules''.  For example, given Colin's
generation and the learned feature of Mother that the output should be one
generation higher than the input, we can already restrict the answer to
people in a specific generation. The ability of deep learning to discover
many such micro-rules allows it to generalize well to novel combinations of
individually familiar context words.

The family tree demonstration used a small, clean dataset in which the
regularities could be captured by micro-rules that were never violated and
the features were easy to interpret, but learned word-vectors also work
very well when the word sequences come from a large corpus of real text and
the individual micro-rules are unreliable\cite{BenDucVin01-short}. When trained to
predict the next word in a news story, for example, the learned word
vectors for Monday and Tuesday are very similar, as are the word vectors
for Vietnam and Iraq.  Vector representations of words learned from text
are now very widely used in natural language applications\cite{Collobert,
  Mikolov, Socher?}.

The issue of representation lies at the heart of the debate between the
logic-inspired and the neurally-inspired paradigms for cognition. In the
logic-inspired paradigm, an instance of a symbol is something whose only
property is that it is either identical or non-identical to other symbol
instances. It has no internal structure that is relevant to its use. By
contrast, neural networks use large activity vectors instead of symbols and
parallel matrix operations followed by non-linearities instead of formal
rules of inference. With sufficient data, computational power and depth,
the combination of backpropagation for computing gradients and stochastic
gradient descent for updating the weights on connections has proved to be
very good at learning appropriate distributed representations. These
representations allow a lot of parallel computation at the feature level so
that complex tasks can be solved in a very small number of sequential
steps. In fact, it can be shown that a measure of the richness of the functions
that can be represented by a deep neural network grows exponentially
with depth~\citep{Montufar-et-al-NIPS2014}.

@@

Recurrent neural networks are very powerful dynamical systems, but training
them with BPTT proved to be problematic, because the backpropagated
gradients either grow or shrink at each time step, so over many time-steps
they typically explode or vanish~\cite{Bengio-trnn93-small,Hochreiter91-small}.

@@

When deep convolutional networks were applied to a dataset of about a
million images from the web that contained a thousand different classes
they achieved spectacular results, almost halving the error rates of the
best computer vision sytems~\cite{kriz}.  Some of the images and the
network's predictions are shown in figure ??.  The major factors in this
success were the large labeled dataset~\cite{FeiFe}, the use of fast GPUs,
and the convolutional early layers.  It had already been shown that deep
supervised nets with rectified linear units could be trained without
unsupervised pre-training~\citep{Glorot+al-AI-2011-small}. 
These rectified linear units made the
training much faster and the fully connected final layers were prevented
from overfitting the peculiarities of the training set by randomly omitting
hidden units during training~\cite{dropoutJMLR}. Deep convolutional nets
have quickly replaced the carefully engineered pipelines full of
hand-designed features developed by the computer vision community and they
are now deployed in many computer vision applications, including image
tagging, object detection, face recognition, facial expression recognition
and scene parsing.

@@

\begin{figure}[H]
\centerline{\includegraphics[width=0.7\linewidth]{word-embeddings.png}}
\caption{Illustration of word representations learned for modeling
language, non-linearly projected to 2-D for visualization.}
\end{figure}

When translating between languages, the hidden units of the decoder network
that generates possible translations are initalized with a ``thought
vector'' which is the final hidden state of an encoder network.  But it is
also possible to initialize a decoder network using a linear transformation
of the last hidden layer of a deep convolutonal network that has already
been trained to recognize objects in images. The decoder network can then
be trained by BPTT on a dataset of captioned
images~\cite{NIC,StanfordNIC,more?} and it produces remarkably good
captions (see figure ??).

@@

\section{Key references}
{\it Quite a few of these are very recent and have not yet been published
  through a peer-review so we are giving their arxiv references}.

\subsection{Back-propagation and Stochastic Gradient Descent}
Rumelhart, D. E., Hinton, G. E., \& Williams, R. J. (1986)\\ Learning
representations by back-propagating errors.\\ {\it Nature}, {\bf 323},
533-536.

LeCun, Y., Boser, B., Denker, J. S., Henderson, D., Howard, R.E., Hubbard,
W., \& Jackel, L.D.\\ Backpropagation Applied to Handwritten Zip Code
Recognition\\ Neural Computation, 1(4):541-551, Winter 1989

\subsection{Distributed representations}
Hinton, G. E., McClelland, J. L., \& Rumelhart, D. E. (1986)\\ Distributed
representations.\\ In Rumelhart, D. E. and McClelland, J. L., editors, {\it
  Parallel Distributed Processing: Explorations in the Microstructure of
  Cognition. Volume 1: Foundations}, MIT Press, Cambridge, MA. pp 77-109.

Bengio, Y., Ducharme, R., \& Vincent, P. (2001)\\ A neural probabilistic
language model.\\ In {\it NIPS 2000}, pages 932--938.

%% Collobert, R., Weston, J., Bottou, L., Karlen, M., Kavukcuoglu, K., Kuksa, P. (2011)\\
%% Natural language processing (almost) from scratch.\\
%% {\it The Journal of Machine Learning Research} 12, pages 2493-2537

Mikolov, T., Sutskever, I., Chen, K., Corrado, G.S., \& Dean
J.\\ Distributed representations of words and phrases and their
compositionality.\\ In {\it NIPS 2013}, pages 3111-3119.

Montufar, G.~F., Pascanu, R., Cho, K., and Bengio, Y. (2014).  On the
number of linear regions of deep neural networks.\\ In {\it NIPS'2014}.

\subsection{The apparent limitations of backpropagation in deep networks}

Bengio, Y., Simard, P., \& Frasconi, P. (1994)\\ Learning long-term
dependencies with gradient descent is difficult.\\ {\it IEEE Transactions
  on Neural Networks}, 5(2), 157-166.

>>>>

So far, most applications of deep learning use feed-forward neural
networks, but for tasks that involve sequential inputs it is often better
to use recurrent neural networks. These are given one element from an input % YB term-->element
sequence at each time step and their recurrent internal ``hidden'' state
depends on the whole history of the input sequence. Recurrent networks are
very good at predicting the next character in a string\cite{Sutskever} or
the next word in a sequence~\citep{Mikolov-et-al-NIPS2013} but they can also be used for
more complex tasks.  After reading an English sentence one word at a time,
for example, an English``encoder'' network can be trained so that the final
state vector of its hidden neurons is a good representation of the thought
expressed by the sentence.  This thought vector can then be used as the
initial hidden state of a trained French ``decoder'' network which outputs
a probability distribution for the first word of the French translation. If
a particular first word is chosen (e.g., according to that probability)
and provided as input to the decoder
network it will then output a probability distribution for the second word
of the translation and so on until a full stop is chosen~\cite{Vinyals,
  Montreal}. 
%YB added a clarification
Overall, this process generates sequences of French words according to
a probability distribution that depends on the English sentence.

@@

The hidden layers of a multi-layer neural network learn to represent the
network's inputs in a way that makes it easy to predict the target
outputs. This is nicely demonstrated by training a multi-layer neural
network to predict the next word in a sequence from a local context of
earlier words~\citep{BenDucVin01-short}. 
Each context word is presented to the network as a one-of-N
vector in which one component has a value of $1$ and the rest are $0$. The
output layer has one unit per word and is forced to produce a probability
distribution over words by using the ``softmax'' function:
\begin{equation}
p_j = \frac{exp(z_j)}{\sum_i exp(z_i)}
\end{equation}
where $z_j$ is the total input received by output unit $j$. The hidden
layers use either the logistic or tanh non-linearity:
\begin{equation}
y_j = \frac{1}{1+ \exp(z_j)}, \ \ \ y_j = \frac{}{} % YB: what was the intent here for the 2nd part?
\end{equation}

The connectivity of the network is restricted so that each of the one-of-N
input vectors must first be converted to a ``word-vector'' before being
used by subsequent layers to give a high probability to the target next
word.  The network learns word-vectors that contain many active components
each of which can be interpreted as a separate feature of the word. This
was first demonstrated\cite{RHW} using sequences of three words, such as
(Colin, Mother, Victoria), derived from two family trees. The learned word
vector for Colin contained ``semantic'' features representing which family
tree Colin was in, what branch of that tree he was in, and from what generation
he was~\cite{Hinton-et-al-PDP1986}. These semantic features were not explicitly present
in the input.  They were discovered by the learning procedure as a good way
of factorizing the structured relationships between the input and output
symbols into multiple ``micro-rules''.  For example, given Colin's
generation and the learned feature of Mother that the output should be one
generation higher than the input, we can already restrict the answer to
people in a specific generation. The ability of deep learning to discover
many such micro-rules allows it to generalize well to novel combinations of
individually familiar context words.

The family tree demonstration used a small, clean dataset in which the
regularities could be captured by micro-rules that were never violated and
the features were easy to interpret, but learned word-vectors also work
very well when the word sequences come from a large corpus of real text and
the individual micro-rules are unreliable~\citep{BenDucVin01-short}. When trained to
predict the next word in a news story, for example, the learned word
vectors for Monday and Tuesday are very similar, as are the word vectors
for Vietnam and Iraq.  Vector representations of words learned from text
are now very widely used in natural language 
applications~\citep{Schwenk-2007,collobert:2011b,Socher-2011,Mikolov-et-al-NIPS2013}.

The issue of representation lies at the heart of the debate between the
logic-inspired and the neurally-inspired paradigms for cognition. In the
logic-inspired paradigm, an instance of a symbol is something whose only
property is that it is either identical or non-identical to other symbol
instances. It has no internal structure that is relevant to its use. By
contrast, neural networks use large activity vectors instead of symbols and
parallel matrix operations followed by non-linearities instead of formal
rules of inference. The elements of such vectors can be understood as
the values of learned attributes (or features) associated with some object. 
With sufficient data, computational power and depth,
the combination of backpropagation for computing gradients and stochastic
gradient descent for updating the weights on connections has proved to be
very good at learning appropriate distributed representations. These
representations allow a lot of parallel computation at the feature level so
that complex tasks can be solved in a very small number of sequential
steps. In fact, it can be shown that a measure of the richness of the functions
that can be represented by a deep neural network grows exponentially
with depth~\citep{Montufar-et-al-NIPS2014}.

@@

Recurrent neural networks are very powerful dynamical systems, but training
them with BPTT proved to be problematic, because the backpropagated
gradients either grow or shrink at each time step, so over many time-steps
they typically explode or vanish~\cite{Bengio-et-al-TNN1994,Hochreiter91-small}.

@@

When deep convolutional networks were applied to a dataset of about a
million images from the web that contained a thousand different classes
they achieved spectacular results, almost halving the error rates of the
best computer vision sytems~\cite{kriz}.  Some of the images and the
network's predictions are shown in figure ??.  The major factors in this
success were the large labeled dataset~\cite{imagenet_cvpr09}, the use of fast GPUs,
and the convolutional early layers.  It had already been shown that deep
supervised nets with rectified linear units could be trained without
unsupervised pre-training~\citep{Glorot+al-AI-2011-small}. 
These rectified linear units made the
training much faster and the fully connected final layers were prevented
from overfitting the peculiarities of the training set by randomly omitting
hidden units during training~\cite{Srivastava14}. Deep convolutional nets
have quickly replaced the carefully engineered pipelines full of
hand-designed features developed by the computer vision community and they
are now deployed in many computer vision applications, including image
tagging, object detection, face recognition, facial expression recognition
and scene parsing.

@@

Prior to the introduction of neural language
models~\citet{BenDucVin01-short} the standard approach to statistical
modeling of language was based on counting frequencies of occurences of
short symbol sequences (called n-grams). One disadvantage is that it
taking into account context beyond a handful of words would require
training corpora of exponentially growing size (with respect to the
desired context size), since the number of configurations of $n$
words from a vocabulary of size $V$ grows as $V^n$. Related to this
problem and the very notion of symbol introduced above, n-grams
do not generalize across the many semantically related sequences 
of words (for example the many ways of expressing roughly the same thing).
Instead, neural language models learn to associate each word to a
distributed representation (a vector) such that semantically related
words end up close to each other in that space, as illustrated in
Figure~\ref{fig:word-embeddings}. To obtain that effect, one
{\em jointly learns} a representation for each word as well as
a function that computes a quantity of interest that is relevant
to a language related task, such as predicting the next word
in a sequence (neural language models) or predicting the sequence
of translated words given the sequence of source words, i.e., neural
machine translation~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.

\begin{figure}[H]
\centerline{\includegraphics[width=0.7\linewidth]{word-embeddings.png}}
\caption{Illustration of word representations learned for modeling
language, non-linearly projected to 2-D for visualization using the
t-SNE algorithm~\citep{VanDerMaaten08}.}
\label{fig:word-embeddings}
\end{figure}

When translating between languages, the hidden units of the decoder network
that generates possible translations are initalized with a ``thought
vector'' which is the final hidden state of an encoder network.  But it is
also possible to initialize a decoder network using a linear transformation
of the last hidden layer of a deep convolutonal network that has already
been trained to recognize objects in images. The decoder network can then
be trained by BPTT on a dataset of captioned
images~\cite{NIC,StanfordNIC,more?} and it produces remarkably good
captions (see Figure~\ref{fig:caption-generation}).

\begin{figure}[ht]
%\centerline{\includegraphics[width=0.7\linewidth]{}}
\caption{Example of captions generated by a recurrent network taking
as extra input the representation extracted by a deep convolutional
net from an image, with the whole system trained to ``translate''
images into captions. Reproduced from~\citet{?}.
}
\label{fig:caption-generation}
\end{figure}

@@

%\section{Key references}
%{\it Quite a few of these are very recent and have not yet been published
%  through a peer-review so we are giving their arxiv references}.

%\subsection{Back-propagation and Stochastic Gradient Descent}

%Rumelhart, D. E., Hinton, G. E., \& Williams, R. J. (1986)\\ Learning
%representations by back-propagating errors.\\ {\it Nature}, {\bf 323},
%533-536.
% --> \citep{Rumelhart86b}

%LeCun, Y., Boser, B., Denker, J. S., Henderson, D., Howard, R.E., Hubbard,
%W., \& Jackel, L.D.\\ Backpropagation Applied to Handwritten Zip Code
%Recognition\\ Neural Computation, 1(4):541-551, Winter 1989
%--> \citep{LeCun89-small}

%\subsection{Distributed representations}

%Hinton, G. E., McClelland, J. L., \& Rumelhart, D. E. (1986)\\ Distributed
%representations.\\ In Rumelhart, D. E. and McClelland, J. L., editors, {\it
%  Parallel Distributed Processing: Explorations in the Microstructure of
%  Cognition. Volume 1: Foundations}, MIT Press, Cambridge, MA. pp 77-109.
% --> \citep{Hinton-et-al-PDP1986}

%Bengio, Y., Ducharme, R., \& Vincent, P. (2001)\\ A neural probabilistic
%language model.\\ In {\it NIPS 2000}, pages 932--938.
% --> \citep{BenDucVin01-short}

%% Collobert, R., Weston, J., Bottou, L., Karlen, M., Kavukcuoglu, K., Kuksa, P. (2011)\\
%% Natural language processing (almost) from scratch.\\
%% {\it The Journal of Machine Learning Research} 12, pages 2493-2537
% --> \citep{collobert:2011b}

%Mikolov, T., Sutskever, I., Chen, K., Corrado, G.S., \& Dean
%J.\\ Distributed representations of words and phrases and their
%compositionality.\\ In {\it NIPS 2013}, pages 3111-3119.
% --> \citep{Mikolov-et-al-NIPS2013}

%Montufar, G.~F., Pascanu, R., Cho, K., and Bengio, Y. (2014).  On the
%number of linear regions of deep neural networks.\\ In {\it NIPS'2014}.
% --> \citep{Montufar-et-al-NIPS2014}.

%\subsection{The apparent limitations of backpropagation in deep networks}

%Bengio, Y., Simard, P., \& Frasconi, P. (1994)\\ Learning long-term
%dependencies with gradient descent is difficult.\\ {\it IEEE Transactions
%  on Neural Networks}, 5(2), 157-166.
% --> \citep{Bengio-et-al-TRNN1994}
